Combinatorial Identities from Lagrange's Interpolation Polynomial

### 2020-12-15 spectral radius of matrix new method/thesis/article121420.tex
121620 update

\documentclass{article}

\topmargin 0pt

\topmargin=-1.5cm
\oddsidemargin=0.7cm
\textheight=23.5cm
\textwidth=15cm
\setlength{\baselineskip}{24pt}
\renewcommand{\baselinestretch}{1.5} %行距

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tabularx,array}
\usepackage{pgf,tikz}
 \usepackage{blkarray} %line 80-92 Q formula matrix index
%\usepackage{colortbl}
\usepackage{enumitem}% enumerate labels   roman
\usepackage{graphicx}
%\usepackage{indentfirst}
\graphicspath{{images/}}
\usetikzlibrary{arrows}
\setCJKmainfont{標楷體} %設定中文為系統上的字型，而英文不去更動，使用原TeX字型
\XeTeXlinebreaklocale "zh"             %這兩行一定要加，中文才能自動換行
\XeTeXlinebreakskip = 0pt plus1pt     %這兩行一定要加，中文才能自動換行
\title{Combinatorial Identities from Lagrange's Interpolation Polynomial}
\author{Student: Yen-Jung Huang  ~~~~~~~~~~~~~~~~~~~~~~~~~~Advisor: Chih-Wen Weng}
\date{} %不要日期

@@

{ \Huge A simple method on estimating spectral radius for some direted graphs}~\\~\\
~\\~\\~\\~\\~\\
\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip
\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip
\bigskip\bigskip\bigskip

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\baselinestretch}{2} %行距
\thispagestyle{empty}
\begin{center}
{
\Large
有向圖譜半徑之簡易估算方法\\
A simple method on estimating spectral radius for some direted graphs\\~\\
\begin{tabular}{lccr}
Student: Ko-Han Chen  &&~~~& Advisor: Chih-Wen Weng\\
研究生:陳科翰　 &&~~~& 指導教授：翁志文~教授
\end{tabular}
}~\\

@@

 我知道譜半徑是區分圖中點連線關係的重要指標，所以找一個簡單的方法來估算譜半徑
 是很有意義的。一個簡單而且傑出估算譜半徑的方法有一些特徵。首先，誤差被極小化
 ，第二必定存在一個方法正確理性地證明它。列舉這些方法並正確證明會讓這方法穩固
 且可信靠。在估計的領域裡，最好思考過所有的因素，再根據目標，取用估計譜半徑的
 方法，以利之後分析圖的連接矩陣，與其連通性有關的矩陣。
 \\
\bigskip

\noindent 關鍵詞：譜半徑，簡易方法，有向圖，估算。
\pagebreak

%英文摘要
\begin{center}{\LARGE
A simple method on estimating spectral radius for some direted graphs
\bigskip\bigskip\bigskip}

@@

%\begin{abstract}
\addcontentsline{toc}{section}{Abstract (in English)}
I know that spectral radius is an important index to specify the relation of connected vertices in
 the graphs, so it is meanful to find a simple method to estimate spectral readius. A simple and excellent
 executable method to estimate spectral radius has some features. First, the bios is minimized, and second
 there must be a way to prove it sensible. Enumerate these factors and prove it correctly would make this method
 solid and reliable. In the field of estimation, it's better to go through all the factors, and base on one's goal
 to apply the methods of estimating spectral radius before analyzing the adjacency matrix of the graphs, which is
 related to the connectness of the graph.

@@

\noindent {\bf Keywords}: spectral radius, simple method, directed graphs, estimation
%\end{abstract}
\pagebreak

@@

\pagebreak

\label{Introduction}
% Introduction

\section{Introduction}
\normalsize
\pagenumbering{arabic}

Let $\mathbb{R}$ and $\mathbb{C}$ denote the
 field of real numbers and complex numbers respectively.

\begin{defn}       
Let $C$ be an $n \times n$ real nonnegative matrix, and $u \in \mathbb{R}^n$ be a
 nonzero column vector. The scalar $\lambda \in \mathbb{C}$ is an $\it{eigenvalue}$ 
of C corresponding to the $\it{eigenvector}$ $u,$  if $Cu = \lambda u.$
%Also, $C$ has an left eigenvector $v^T$, when $v^TC=\lambda v^T$.
\end{defn}

\begin{defn} \cite{spec_rad}
    When $C$ is an $n \times n$ real matrix, the $\textit {spectral radius} $ $\rho(C)$
        of $C$ is defined by 
        $$\rho(C):=\max\{~|\lambda|~\ |~~\lambda\hbox{ is an eigenvalue of $C$}\},$$
    where $|\lambda|$ is the magnitude of complex number $\lambda.$
\end{defn}

We are interested in spectral radius of the following matrix associated with a simple graph.

\begin{defn} 
    Given an directed graph G, the$\textit{ adjacency matrix}$ of G is the square 
    matrix $A = (a_{ij})$ indexed by vertices of G, and
     \[a_{ij} =\begin{cases} 
        1, \text{if $i$ is adjacent to $j$}, \\
        0, \text{otherwise.}
            \end{cases}
     \]
\end{defn}

\begin{defn}
    Given a directed graph G, the $\textit{spectral radius}$  $\rho(G) $ of G is the spectral
     radius of the adjacency matrix of G.
\end{defn}

\section{Preliminaries}

The next theorem is Perron–Frobenius Theorem, which provides a feature of
 nonnegative eigenvector to nonnegative matrices.
\begin{thm} \cite{chang} \cite{prn_fros2} \label{thm:Perron_Frobenius}
    If $C$ is nonnegative square matrix, then the spectral radius $\rho(C)$ is an
    eigenvalue of $C$ with a corresponding nonnegative right eigenvector and a
    corresponding nonnegative left eigenvector.
\end{thm}

@@

\begin{defn}
    For a matrix $C=(c_{ij})$ and subsets $\alpha$, $\beta$ of row indices and column 
    indices of $C$ respectively,  We use $C[\alpha|\beta]$ to denote the 
    submatrix of $C$ with size $ |\alpha| \times |\beta| $ that has entries $c_{ij}$ for $i\in \alpha$
    and $j\in\beta$,
\end{defn}

We introduce two matrices $P$ and $Q$ in the following Theorem, where $P$ is a permutation
 matrix which is multiplied to the left side, and $Q$ is sum of elementary matrix and
 certain binary matrix. In which $P$ generalize row permutation on cases of $C$
 matrix,and $Q$ is the transform from $C'$ to C', which is the first n-1 columns
  and sum of certain columns. We aim to find C' such that $C'$ majors $C$,i.e. $C\leq C'$

The following Theorem is from \cite{chang}.

@@

\begin{proof}
    Multiplying the nonnegative vector $u$ in Theorem~\ref{pre_thm} assumption
    ~\ref{pre_thm_em1}, where $Qu$ is eigenvector of $C'$,  to the right of both terms of
    ~\ref{pre_thm_em1},    
    \begin{equation}\label{pre1}
       PCQu\leq PC'Qu=\lambda'PQu.
    \end{equation}
    Multiplying the nonnegative left eigenvector $v^T$ of $C$ for $\lambda$ in assumption
     ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}), where $v^TP$ is
    left eigenvector of $C$ for $\lambda$, thus we have
    \begin{equation}\label{pre2}
        \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
    \end{equation}
        Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
        $\lambda\leq \lambda'$ and finish the proof of the first part.
        Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
        Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
        $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
        $u_j\not=0.$ Conversely, (\ref{pre0}) implies $$v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j=
         \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so $\lambda=\lambda'$ by (\ref{pre2}).
\end{proof}

\section{Our Method}
We use $[n-1]$ as notation of the set of elements from one to $n-1$, which is ${1,2,...,n}$.
Throughout fix $k\in [n-1]$.Let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$ 
appearing in the position $k,n$ of $E_{kn}$. We will apply the previous Theorem \ref{pre_thm} with $P=I$ and

\begin{equation} \label{Q_1}
Q=I+E_{kn}=\begin{pmatrix} 
1 &  & & &  & 0 \\
 & 1 &  &      &  &  \\
 &  & \ddots & &  & 1 \\
 &  &        & &  &  \\
  &  & & & 1 &  \\
0 &  & & &  & 1 \\
\end{pmatrix}.  
\end{equation}

\begin{defn}%[$K$-rooted vector]
~A column vector $v'=(v'_1,v'_2,\ldots,v'_n)^T$ is called {\it $K$-rooted}  if $v'_{j} \geq 0$ for $1 \leq  j \leq n$ and $v'_k\geq v'_n.$
\end{defn}
\bigskip

The following Lemma is immediate from the above definition.%[vector rooted lemma]
\bigskip

\begin{lem}\label{lem:rt_vec}
If $u=(u_1, u_2, \ldots, u_n)^T$ and $v'=(v'_1, v'_2, \ldots, v'_n):=Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$, then
\begin{enumerate}[label=(\roman*)]
\item \label{lem:rt_vec:en1}$v'$ is $K$-rooted  if and only if  $u$ is nonnegative;
\item $u_k>0$ if and only if $v'_k>v'_n$.
\end{enumerate}
% \qed
\end{lem}

Below is our first result, in which the first condition implies the first n-1 columns of C major to
 the columns of C', and the (k,n)-sum column of C is also major to C'. The second and the third condition
 suggest that C and C' have nonnegative eigenvectors which are k-rooted. And the forth condition is simplier
  but with the same meaning with Theorem (\ref{pre_thm})
\begin{thm}\label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\roman*)]
\item \label{thm_main:condition_i} $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item \label{thm_main:condition_ii} there exists a $K$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item \label{thm_main:condition_iii}there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item \label{thm_main:condition_iv}$v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}[label=(\alph*)]
    \item \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
    \item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $with $v_i\ne 0 $;
    \item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$ 
\end{enumerate} %\qed
\end{thm}
%

\begin{proof}
    The proof is based on Theorem~\ref{pre_thm} with $P = I$ and $Q = I + E_{kn}$ in (\ref{Q_1}). 
    The assumption \ref{pre_thm_em1} $PCQ\leq PC'Q$ of Theorem~\ref{pre_thm} holds by the condition \ref{thm_main:condition_i} of this Theorem. 
    Let $u = Q^{-1}v'$. Then u is nonnegative and $C'Qu = \lambda' Qu$ by the condition \ref{thm_main:condition_ii} and
     Lemma~\ref{lem:rt_vec}\ref{lem:rt_vec:en1}. Hence the assumption \ref{pre_thm_em2} of Theorem~\ref{pre_thm} holds. The assumptions \ref{pre_thm_em3} and \ref{pre_thm_em4}
      of Theorem~\ref{pre_thm} clearly hold by conditions~\ref{thm_main:condition_iii},\ref{thm_main:condition_iv} of this Theorem since $P = I$ and
       $v'= Qu$  Hence $\lambda \leq \lambda' $ by the necessary condition of Theorem~\ref{pre_thm}. Moreover
        $\lambda = \lambda'$ if and only if \ref{pre0} holds, and this is equivalent to
         conditions \ref{thm_main:equ_cond_a},\ref{thm_main:equ_cond_b},\ref{thm_main:equ_cond_c} of this Theorem. 
\end{proof}

We are interested in the matrices $C'$ that have $K$-rooted eigenvectors.
Motivated by the condition (i) of Theorem 2.3, we provide the following two definitions. 
This is the definition of (k,n)-sum.
\begin{defn}
    For an $n \times n$ matrix $C'=(c'_{ij})$, the $(k, n)$-sum vector of $C'$ is the vector of the sum of the $k$-th and  $n$-th columns of $C'$, where $k\leq n-1$.
\end{defn}

Note that the last column of $C'Q$ is the $(k, n)$-sum vector of $C'$
Below is the definition of k-rooted matrix.
\begin{defn}\label{m_rooted}
    A  matrix $C'=(c'_{ij})$ is called $k-rooted$  if its  columns and its $(k, n)$-sum vector are all $K$-rooted except the last column of $C'$.
\end{defn}

@@

    $$Q^{-1}=I-E_{kn}=\begin{pmatrix}
    1 &  & & &  & 0 \\
    & 1 &  &      &  &  \\
    &  & \ddots & &  & -1 \\
    &  &        & &  &  \\
    &  & & & 1 &  \\
    0 &  & & &  & 1 \\
    \end{pmatrix}.$$

    The matrix $C'Q$ is
    $$\begin{pmatrix}
    c'_{11}     & c'_{12} & \cdots     & c'_{1\ n-1} & c'_{1k}+c'_{1n} \\
    \vdots \\
    c'_{k-11}     & c'_{k-1 2}           & \cdots     & c'_{k-1 n-1} & c'_{k-1k}+c'_{k-1n} \\
    c'_{k1} & c'_{k2} &\cdots      & c'_{kn-1} & c'_{kk}+c'_{kn}\\
    c'_{k+11}     & c'_{k+12}           & \cdots     & c'_{k+1\ n-1} & c'_{k+1k}+c'_{k+1n} \\
    \vdots              & \vdots & \ddots              & \vdots & \vdots \\
    c'_{n1}             & c'_{n2} & \cdots             & c'_{n\ n-1} & c'_{nk}+c'_{nn} \\
    \end{pmatrix}.
    $$
\end{rem}

The following Theorem shows that a $K$-rooted matrix has a $K$-rooted eigenvector.
\begin{lem}\label{lma_m_rooted}
    Let $C'=(c'_{ij})$ be an $n\times n$ nonnegative matrix. Then the following (i)-(iii) hold.
        \begin{enumerate}[label=(\roman*)]
            \item \label{lma_m_rooted_cond1} $C'$ is a $K$-rooted matrix, if and only if, $Q^{-1}C'Q$ is nonnegative.
            \item \label{lma_m_rooted_cond2}Assume that $C'$ is $K$-rooted and let $u$ be a nonnegative eigenvector of $Q^{-1}C'Q$
                for $\rho(C')$. Then  $C'$ has a $K$-rooted eigenvector $v'=Qu$ for $\rho(C')$. 
            \item \label{lma_m_rooted_cond3}$\rho(C')$ = $\rho(Q^{-1}C'Q)$
        \end{enumerate}
\end{lem}

\begin{comment}

>>>>

\documentclass[12pt, a4paper]{article}

\setlength{\textwidth}{160mm} \setlength{\textheight}{220mm}
\topmargin=-0.0cm \oddsidemargin=-0.1cm \setcounter{page}{1}

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tabularx,array}
\usepackage{pgf,tikz}
 \usepackage{blkarray} %line 80-92 Q formula matrix index
%\usepackage{colortbl}
\usepackage{enumitem}% enumerate labels   roman
\usepackage{graphicx}

\graphicspath{{images/}}
\usetikzlibrary{arrows}
\setCJKmainfont{標楷體} %設定中文為系統上的字型，而英文不去更動，使用原TeX字型
\XeTeXlinebreaklocale "zh"             %這兩行一定要加，中文才能自動換行
\XeTeXlinebreakskip = 0pt plus1pt     %這兩行一定要加，中文才能自動換行
\title{Combinatorial Identities from Lagrange's Interpolation Polynomial}
\author{Student: Yen-Jung Huang  ~~~~~~~~~~~~~~~~~~~~~~~~~~Advisor: Chih-Wen Weng}
\date{} %不要日期

@@

{ \Huge A simple method on estimating spectral radius for some directed graphs}~\\~\\
~\\~\\~\\~\\~\\
\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip
\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip
\bigskip\bigskip\bigskip

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\baselinestretch}{2} %行距
\thispagestyle{empty}
\begin{center}
{
\Large
有向圖譜半徑之簡易估算方法\\
A simple method on estimating spectral radius for some directed graphs\\~\\
\begin{tabular}{lccr}
Student: Ko-Han Chen  &&~~~& Advisor: Chih-Wen Weng\\
研究生:陳科翰　 &&~~~& 指導教授：翁志文~教授
\end{tabular}
}~\\

@@

 矩陣的譜半徑為其特徵值絕對值的最大值，而一有向圖的譜半徑則定義為其鄰接矩陣之譜半徑。本論文給出一個尋找方陣譜半徑上界的方法，將此方法應用於有向圖的鄰接矩陣，我們可以簡易估算出此圖的譜半徑。
 \\
\bigskip

\noindent 關鍵詞：譜半徑、鄰接矩陣
\pagebreak

%英文摘要
\begin{center}{\LARGE
A simple method on estimating spectral radius for some directed graphs
\bigskip\bigskip\bigskip}

@@

%\begin{abstract}
\addcontentsline{toc}{section}{Abstract (in English)}
The spectral radius of a square matrix is the largest magnitude of its eigenvalues. And the spectral radius of a directed graph is defined as the spectral radius of the corresponding adjacency matrix. In this paper, we give an approach to obtain an upper bound for the spectral radius of a square matrix. By applying this method on the adjacency matrix of a directed graph, we can estimate the spectral radius of this directed graph simply.

@@

\noindent {\bf Keywords}: spectral radius, adjacency matrix
%\end{abstract}
\pagebreak

@@

\clearpage
\pagenumbering{arabic}
\linespread{1.5}

\setcounter{page}{1}
%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%

Let $\mathbb{R}$ and $\mathbb{C}$ denote the field of real numbers and complex numbers, respectively. Let $C$ be an $n\times n$ real square matrix. If there is a nonzero column vector $u\in\mathbb{R}^n$ such that $Cu=\lambda u$ for some scalar $\lambda\in\mathbb{C}$, then the scalar $\lambda$ is called the eigenvalue of $C$ with corresponding eigenvector $u$. And the spectral radius of a matrix $C$ is the largest magnitude (or complex modulus) of its eigenvalues, denoted by $\rho(C)$. We are interested in the spectral radius of the following matrix associated with a simple directed graph.

\begin{defn}
    Given an directed graph $G$, the $\textit{adjacency matrix}$ of $G$ is the square
    matrix $A = (a_{ij})$ indexed by vertices of $G$, and
     \[a_{ij} =\begin{cases}
        1, \text{if $i$ is adjacent to $j$}, \\
        0, \text{otherwise.}
            \end{cases}
     \]
\end{defn}

Given a directed graph $G$, the spectral radius of $G$ is the spectral radius of the adjacency matrix of $G$, denoted by $\rho(G)$. Note that the spectral radius $\rho(G)$ is independent of the ordering of the vertex set of $G$.

The spectral radius is an important indicator to specify the relation of connected vertices in a graph, so it is meaningful to find a simple method to estimate the spectral radius. A simple and excellent executable method to estimate the spectral radius has some features, first, the bios is minimized, and second, there must be a way to prove it sensible. Enumerate these factors and prove it correctly would make this method reliable.

In \cite{chang}, Cheng and Weng give many bounds of the spectral radius of a nonnegative square matrix. And based on their theory and Perron-Frobenius theorem, we give another approach to obtain an upper bound of the spectral radius and apply it on the adjacency matrix of a directed graph.

%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminaries}
%%%%%%%%%%%%%%%%%%%%%%%

The following is Perron–Frobenius theorem, which provides a feature of
 nonnegative eigenvector to nonnegative matrices.

\begin{thm} \cite{prn_fros2} \label{thm:Perron_Frobenius}
    If $C$ is a nonnegative square matrix, then the spectral radius $\rho(C)$ is an
    eigenvalue of $C$ with a corresponding nonnegative right eigenvector and a
    corresponding nonnegative left eigenvector.
\end{thm}

@@

\begin{defn}
    For a matrix $C=(c_{ij})$ and subsets $\alpha$, $\beta$ of row indices and column
    indices of $C$, respectively, we use $C[\alpha|\beta]$ to denote the
    submatrix of $C$ with size $ |\alpha| \times |\beta| $ that has entries $c_{ij}$ for $i\in \alpha$
    and $j\in\beta$.
\end{defn}

Our theory is based on the following theorem, which is from \cite{chang}, and we introduce two square matrices $P$ and $Q$ in the following theorem first. The matrix $P$ is a permutation matrix which is multiplied to the left side, and the matrix $Q$ is the sum of an elementary matrix and a certain binary matrix. In which $P$ generalize row permutation on cases of $C$ matrix, and $Q$ is the transform from $C'$ to $C'$, which is the first $n-1$ columns and sum of certain columns. We aim to find $C'$ such that $C'$ majors $C$, i.e. $C\leq C'$.

@@

\begin{proof}
    Multiplying the nonnegative vector $u$ in assumption~\ref{pre_thm_em2}, where $Qu$ is eigenvector of $C'$, to the right of both terms of
    ~\ref{pre_thm_em1},
    \begin{equation}\label{pre1}
       PCQu\leq PC'Qu=\lambda'PQu.
    \end{equation}
    Multiplying the nonnegative left eigenvector $v^T$ of $C$ for $\lambda$ in assumption
     ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}), where $v^TP$ is
    left eigenvector of $C$ for $\lambda$, thus we have
    \begin{equation}\label{pre2}
        \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
    \end{equation}
        Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
        $\lambda\leq \lambda'$ and finish the proof of the first part.
        Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
        Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
        $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
        $u_j\not=0.$ Conversely, (\ref{pre0}) implies $$v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j=
         \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so $\lambda=\lambda'$ by (\ref{pre2}).
\end{proof}

%%%%%%%%%%%%%%%%%%%%
\section{Our Method}
%%%%%%%%%%%%%%%%%%%%

We use $[n-1]$ as notation of the set of elements from one to $n-1$, which is $\{1,2,...,n-1\}$.
Throughout fix $k\in [n-1]$, let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$
appearing in the position $k,n$ of $E_{kn}$. Now we apply the previous Theorem \ref{pre_thm} with $P=I$ and

\begin{equation} \label{Q_1}
Q=I+E_{kn}=\begin{pmatrix}
1 &   &         &   & 0 \\
  & 1 &         &   &   \\
  &   & \ddots  &   & 1 \\
  &   &         & 1 &  \\
0 &   &         &     & 1
\end{pmatrix},
\end{equation}
so the matrix $PC'Q$ in assumption~\ref{pre_thm_em1} of Theorem~\ref{pre_thm} is

\begin{equation}\label{PC'Q}
  PC'Q=\begin{pmatrix}
         c'_{11} & c'_{12} & \cdots &  c'_{1k}+c'_{1n} \\
         c'_{21} & c'_{22} & \cdots &  c'_{2k}+c'_{2n} \\
         \vdots & \vdots & \ddots &  \vdots \\
         c'_{n1} & c'_{n2} & \cdots &  c'_{nk}+c'_{nn}
       \end{pmatrix},
\end{equation}
where $c'_{ij}$ denote the $(i,j)$-entry of $C$.

\begin{defn}%[$K$-rooted vector]
A column vector $v'=(v'_1,v'_2,\ldots,v'_n)^T$ is called {\it $K$-rooted}  if $v'_{j} \geq 0$ for $1 \leq  j \leq n$ and $v'_k\geq v'_n.$
\end{defn}

The following Lemma is immediate from the above definition.%[vector rooted lemma]

\begin{lem}\label{lem:rt_vec}
If $u=(u_1, u_2, \ldots, u_n)^T$ and $v'=(v'_1, v'_2, \ldots, v'_n):=Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$, then
\begin{enumerate}[label=(\roman*)]
\item \label{lem:rt_vec:en1}$v'$ is $k$-rooted  if and only if $u$ is nonnegative;
\item $u_k>0$ if and only if $v'_k>v'_n$.
\end{enumerate}
% \qed
\end{lem}

Below is our first result, in which the first condition implies the first $n-1$ columns of $C$ major to
 the columns of $C'$, and the $(k,n)$-sum column of $C$ is also major to $C'$. The second and the third condition
 suggest that $C$ and $C'$ have nonnegative eigenvectors which are $K$-rooted. And the forth condition is simpler
  but with the same meaning with Theorem~\ref{pre_thm}
\begin{thm}\label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\roman*)]
\item \label{thm_main:condition_i} $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item \label{thm_main:condition_ii} there exists a $K$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item \label{thm_main:condition_iii}there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item \label{thm_main:condition_iv}$v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}[label=(\alph*)]
    \item \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
    \item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $with $v_i\ne 0 $;
    \item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$
\end{enumerate} %\qed
\end{thm}
%

\begin{proof}
    The proof is based on Theorem~\ref{pre_thm} with $P = I$ and $Q = I + E_{kn}$ in (\ref{Q_1}).
    The assumption \ref{pre_thm_em1} $PCQ\leq PC'Q$ of Theorem~\ref{pre_thm} holds by the condition \ref{thm_main:condition_i} of this theorem.
    Let $u = Q^{-1}v'$. Then u is nonnegative and $C'Qu = \lambda' Qu$ by the condition \ref{thm_main:condition_ii} and \ref{lem:rt_vec:en1} in
     Lemma~\ref{lem:rt_vec}. Hence the assumption \ref{pre_thm_em2} of Theorem~\ref{pre_thm} holds. The assumptions \ref{pre_thm_em3} and \ref{pre_thm_em4}
      of Theorem~\ref{pre_thm} clearly hold by conditions~\ref{thm_main:condition_iii}, \ref{thm_main:condition_iv} of this theorem since $P = I$ and
       $v'= Qu$. Hence $\lambda \leq \lambda' $ by the necessary condition of Theorem~\ref{pre_thm}. Moreover,
        $\lambda = \lambda'$ if and only if (\ref{pre0}) holds, and this is equivalent to
         conditions \ref{thm_main:equ_cond_a}, \ref{thm_main:equ_cond_b} and \ref{thm_main:equ_cond_c} of this theorem.
\end{proof}

We are interested in the matrices $C'$ that have $K$-rooted eigenvectors.
Motivated by the condition (i) of Theorem 2.3, we provide the following two definitions.
This is the definition of $(k,n)$-sum.
\begin{defn}
    For an $n \times n$ matrix $C'=(c'_{ij})$, the $(k, n)$-sum vector of $C'$ is the vector of the sum of the $k$-th and  $n$-th columns of $C'$, where $k\leq n-1$.
\end{defn}

Note that the last column of $C'Q$ is the $(k, n)$-sum vector of $C'$.
Below is the definition of $K$-rooted matrix.
\begin{defn}\label{m_rooted}
    A  matrix $C'=(c'_{ij})$ is called {\it $K$-rooted} if its first $n-1$ columns and its $(k,n)$-sum vector are all $K$-rooted.
\end{defn}

@@

    $$Q^{-1}=I-E_{kn}=\begin{pmatrix}
    1 &  & &   & 0 \\
    & 1 &        &  &  \\
    &  & \ddots  &  & -1 \\
    &  & &  1 &  \\
    0 &  &  &  & 1
    \end{pmatrix}.$$

\end{rem}

The following lemma shows that a $K$-rooted matrix has a $K$-rooted eigenvector.
\begin{lem}\label{lma_m_rooted}
    Let $C'=(c'_{ij})$ be an $n\times n$ nonnegative matrix. Then the following (i)-(iii) hold.
        \begin{enumerate}[label=(\roman*)]
            \item \label{lma_m_rooted_cond1} $C'$ is a $K$-rooted matrix if and only if $Q^{-1}C'Q$ is nonnegative,
            \item \label{lma_m_rooted_cond2}Assume that $C'$ is $K$-rooted and let $u$ be the nonnegative eigenvector of $Q^{-1}C'Q$
                corresponding to $\rho(C')$. Then  $C'$ has a $K$-rooted eigenvector $v'=Qu$ for $\rho(C')$,
            \item \label{lma_m_rooted_cond3}$\rho(C')$ = $\rho(Q^{-1}C'Q)$.
        \end{enumerate}
\end{lem}

\begin{comment}
\begin{rem}
        (i) is immediate from Definition~\ref{m_rooted} and the observation that
        $$Q^{-1}=I-E_{kn}=\begin{pmatrix}
        1 &  & & &  & 0 \\
         & 1 &  &      &  &  \\
         &  & \ddots & &  & -1 \\
         &  &        & &  &  \\
          &  & & & 1 &  \\
        0 &  & & &  & 1 \\
        \end{pmatrix},$$
\end{rem}
\end{comment}
\begin{proof}
\begin{enumerate}
  \item[(i)] The matrix $Q^{-1}C'Q$ is
        $$\begin{pmatrix}
            c'_{11}     & c'_{12} & \cdots     & c'_{1\ n-1} & c'_{1k}+c'_{1n} \\
            \vdots      & \vdots  & \vdots     & \vdots      & \vdots\\
            c'_{(k-1) 1}   & c'_{(k-1)  2}           & \cdots     & c'_{(k-1) (n-1)} & c'_{(k-1) k}+c'_{(k-1) n} \\
            c'_{k1}-c'_{n1} & c'_{k2}-c'_{n2} &\cdots      &c'_{k (n-1)}-c'_{n (k-1)}& c'_{kk}+c'_{kn}-c'_{nk}-x'_{nn}\\
            c'_{(k+1) 1}   & c'_{(k+1) 2}           & \cdots     & c'_{(k+1) (n-1)} & c'_{(k+1) k}+c'_{(k+1)
             n} \\
            \vdots              & \vdots & \ddots              & \vdots & \vdots \\
            c'_{n1}     & c'_{n2} & \cdots             & c'_{n (n-1)} & c'_{nk}+c'_{nn} \\
        \end{pmatrix},$$

        which is nonnegative if and only if $C'$ is $K$-rooted by the definition of nonnegative matrix and $K$-rooted matrix.
  \item[(ii)] By Lemma~\ref{lem:rt_vec}, $v'=Qu$ is $K$-rooted. Since $Q^{-1}C'Qu=\rho(C')u$ by the assumption, we have
      $$Q^{-1} C' Q u  = Q^{-1} \rho(C') Qu  =\rho(C')u,$$
      which implies
        $$C'Qu=\rho(C')Qu.$$
  \item[(iii)] Since $C'$ and $Q^{-1}C'Q$ have the same set of eigenvalues, clearly
  $$\rho(C') = \rho(Q^{-1}C'Q).$$
\end{enumerate}
\end{proof}

\begin{lem}\label{l_diag}   %end of label
If a square matrix $C'$ has a $K$-rooted eigenvector corresponding to $\lambda'$, then $C'+dI$ also has
the same $K$-rooted eigenvector corresponding to $\lambda'+d,$ where $d$ is a constant and $I$ is the identity matrix with the same size of $C'$.
\end{lem}
